---
title: "Convert CONAGUA Temperatures to Raster File"
output: html_notebook
---

```{r}
library(dplyr)
library(stringr)
library(stringi)
library(tidyverse)
library(raster)
library(maptools)
library(sp)
library(gstat)
```


## Create a list of files that are Mensuales 
Yes, when we downloaded the files, we did not check whether the files names were consistent.
Then we filter on file urls that contains the 'Mensuales' keyword to scrap the monthly average value per station.

```{r}
cols_url <- paste("V", 1:7,sep='_')
df.files <- read.csv('data/conagua_files_list.csv')
df.files.gather<- df.files %>% select(c('ID',cols_url )) %>% gather(key=vector,value=url,-ID)
df.files.mensuales <- df.files.gather %>% dplyr::filter(.,grepl('Mensuales',url))
str(df.files.mensuales)
head(df.files.mensuales)
df.files.mensuales$filename<- paste(df.files.mensuales$ID,df.files.mensuales$vector,'.txt',sep='')
```



## Combine all stations info into dataframe

```{r}
all_files <-list.files(DATA_FOLDER <- 'data/sonaguas')
all_infos_files <- all_files[grepl('(V_1)',all_files)]
all_temp_files <- df.files.mensuales$filename

```


## Read the full history file and retrieve the station informations 

```{r}
get_stations_infos <- function(station_id){
  
  tryCatch({
  
  DATA_FOLDER <- 'data/sonaguas'
  file <- paste(station_id,'V_1.txt',sep='')
  filename <-paste(DATA_FOLDER,file,sep='/')
  
  con <- file(filename)
  lns <- readLines(filename)
  
  start.read.infos <- grep('ESTAC',lns)
  end.read.infos <- grep('EMISI',lns)
  
  df.infos <- read.table(filename, skip=start.read.infos-1,nrows=end.read.infos- start.read.infos-1, sep=':',col.names = c('feature','value'),encoding="Latin-1") 
  
  df.infos <- df.infos %>% mutate_all(funs(gsub("\260",'',.)))
  df.infos$feature <- gsub(' ','',df.infos$feature)
  df.infos$feature <- gsub('\323','O',df.infos$feature)
  df.infos$value <- trimws(df.infos$value)
  df.infos<-df.infos %>% spread(feature,value)
  colnames(df.infos) <- make.names(colnames(df.infos), unique=TRUE)
  
  df.infos$LONGITUD<- as.numeric(df.infos$LONGITUD)
  df.infos$LATITUD<- as.numeric(df.infos$LATITUD)
  
  close(con)
  
  },error=function(cond) {
            message(paste("file could not be wrapped :", station_id))
  })
  return(df.infos)
}

```




```{r}
get.all.stations.infos <- function(){
  for (infos_file in all_infos_files){
  station <- strsplit(infos_file, "V")[[1]]
  station_id <- station[1]
  
  df.infos.stations <-get_stations_infos(station_id)
  
    if (exists('df.all') ){
    df.all <- bind_rows(df.all,df.infos.stations)

  } else{
    df.all <-df.infos.stations
  }
  }
  write.csv(df.all,'data/conagua_stations_infos.csv')
}

#get.all.stations.infos()


```


## Read the temperature text file for a given Station and convert the MIN/MAX Tem Prom to Dataframe

```{r}
convert_prom_temp_to_dataframe <- function(f){
  
  df.tmp.min<- as.data.frame(c(1))
  df.tmp.max<- as.data.frame(c(1))
  
  tryCatch({
 
    DATA_FOLDER <- 'data/sonaguas/'
    
    file_name <-paste(DATA_FOLDER,f,sep='')
    con <- file(file_name)
    lns <- readLines(con)
    
    MINIMA <- grep('MINIMA',lns)
    start.read.mintmp <- grep('TEMP MINIMA PROM',lns)
    start.read.maxtmp <- grep('TEMP MAXIMA PROM',lns)
    
    
    col.names <- c('title', 'YEAR',month.abb[seq(1:12)],'ACCUMULATED','MEAN', 'MONTHS')
    
    if(length(start.read.mintmp)>0){
      
      
      end.read.mintmp <- MINIMA[match(start.read.mintmp,MINIMA)+1]
       tmp.min.rows <- end.read.mintmp - start.read.mintmp -2
       
      df.tmp.min<- read.fwf(file_name, width=c(25,5,rep(7,each=13),10,10),skip=start.read.mintmp,col.names=col.names,n=tmp.min.rows)
    df.tmp.min$title='TEMP MINIMA PROM'
    }
    
     if(length(start.read.maxtmp)>0){
        end.read.maxtmp <- MINIMA[MINIMA>start.read.maxtmp][1]
          tmp.max.rows <- end.read.maxtmp - start.read.maxtmp -2
    df.tmp.max<- read.fwf(file_name, width=c(25,5,rep(7,each=13),10,10),skip=start.read.maxtmp,col.names=col.names,n=tmp.max.rows)
    df.tmp.max$title='TEMP MAXIMA PROM'
     }
    
    df<-  rbind(df.tmp.min,df.tmp.max)
    df$ID <- file_name
    
    df <- df %>%  select(ID, everything())
    close(con)
    
  },error=function(cond) {
            message(paste("file could not be wrapped :", file_name))
  })
    
    return (df)
}



```

### Retrieve temp for stations without V_7 files (use V_3 instead)
```{r}

retrieve.all.temps.to.df <- function(){
  df<-as.data.frame(c(1))
  for (f in all_temp_files){
  df<- bind_rows(df,convert_prom_temp_to_dataframe(f))
  }
  df<- df %>% mutate_at(vars(c(ID)) ,funs(str_extract(.,"[^/]+(?=.txt)")))
 df[1]<-NULL
write.csv(df,'data/conagua_stations_temp.csv')
}

```

### Merge stations infos with stations temperatures
```{r}
df.stations.infos<- read.csv('data/conagua_stations_infos.csv')
df.stations.temp<- read.csv('data/conagua_stations_temp.csv')
df.stations.infos$ESTACION <- as.character(df.stations.infos$ESTACION)
df.stations.infos$X <- NULL
df.stations.temp$ESTACION <- str_extract(df.stations.temp$ID,"[^/]+(?=V)")
df.stations.temp$X<-NULL
df.stations.temp$c.1.<- NULL
df.stations.temp <- df.stations.temp %>% filter(!is.na(ESTACION))
df.stations.infos.temp <- merge(df.stations.temp,df.stations.infos, by='ESTACION')
anti_join(df.stations.temp,df.stations.infos,by='ESTACION') %>% group_by(ID) %>% summarise(n())

```

### Convert the temperatures dataframe into raster file
```{r}
spg.df<- df.stations.infos.temp
spg.df <- spg.df %>% filter (!is.na(ID)) %>%  filter(YEAR==2010) %>% filter(title=='TEMP MINIMA PROM') %>%  filter(!is.na(Jan))%>% dplyr::select(LATITUD,LONGITUD,Jan)
colnames(spg.df)<- c('y','x','Jan')
spg<- spg.df
coordinates(spg) <- ~  x + y 
plot(spg)

x.range <- as.numeric(c(-116.908, -86.82))  # min/max longitude of the interpolation area
y.range <- as.numeric(c(14.618, 32.715))  # min/max latitude of the interpolation area

grd <- expand.grid(x = seq(from = x.range[1], to = x.range[2], by = .1), y = seq(from = y.range[1], 
    to = y.range[2], by = .1))  # expand points to grid
coordinates(grd) <- ~x + y
gridded(grd) <- TRUE
plot(grd, cex = 1.5, col = "grey")
points(spg, pch = 1, col = "red", cex = 1)
```

```{r}
idw <- idw(formula = Jan ~ 1, locations = spg, 
    newdata = grd)  # apply idw model for the data

idw.output = as.data.frame(idw)  # output is defined as a data table
names(idw.output)[1:3] <- c("long", "lat", "temp.pred")  # give names to the modelled variables


```

```{r}
ggplot() + geom_tile(data = idw.output, aes(x = long, y = lat, fill = ((temp.pred))))+geom_point(data = spg.df, aes(x = x, y=y),shape=1,size=.1)+
  labs(title='Conagua Jan 2010 Min Temp Prom')+ scale_fill_distiller(palette = "Spectral")
```

