---
title: "Convert CONAGUA Temperatures to Raster File"
output: html_notebook
---

```{r}
library(dplyr)
library(stringr)
library(stringi)
library(tidyverse)


Sys.setlocale("LC_MESSAGES", 'en_GB.UTF-8')
Sys.setlocale('LC_ALL','en_GB.UTF-8')
Sys.setenv(LANG = "en_US.UTF-8")
```


## Create a list of files that are Mensuales 
Yes, when we downloaded the files, we did not check whether the files names were consistent.
Then we filter on file urls that contains the 'Mensuales' keyword to scrap the monthly average value per station.

```{r}
cols_url <- paste("V", 1:7,sep='_')
df.files <- read.csv('data/conagua_files_list.csv',encoding = 'utf-8')
df.files.gather<- df.files %>% dplyr::select(c('ID',cols_url )) %>% gather(key=vector,value=url,-ID)
df.files.mensuales <- df.files.gather %>% dplyr::filter(.,grepl('Mensuales',url))
str(df.files.mensuales)
head(df.files.mensuales)
df.files.mensuales$filename<- paste(df.files.mensuales$ID,df.files.mensuales$vector,'.txt',sep='')
```



## Combine all stations info into dataframe

```{r}
all_files <-list.files(DATA_FOLDER <- 'data/sonaguas')
all_infos_files <- all_files[grepl('(V_1)',all_files)]
all_temp_files <- df.files.mensuales$filename

```


## Read the full history file and retrieve the station informations 

```{r}
get_stations_infos <- function(station_id){
  
  tryCatch({
  
  DATA_FOLDER <- 'data/sonaguas'
  file <- paste(station_id,'V_1.txt',sep='')
  filename <-paste(DATA_FOLDER,file,sep='/')
  
  con <- file(filename)
  lns <- readLines(filename)
  
  start.read.infos <- grep('ESTAC',lns)
  end.read.infos <- grep('EMISI',lns)
  
  df.infos <- read.table(filename, skip=start.read.infos-1,nrows=end.read.infos- start.read.infos-1, sep=':',col.names = c('feature','value'),encoding="Latin-1") 
  
  df.infos <- df.infos %>% mutate_all(funs(gsub("\260",'',.)))
  df.infos$feature <- gsub(' ','',df.infos$feature)
  df.infos$feature <- gsub('\323','O',df.infos$feature)
  df.infos$value <- trimws(df.infos$value)
  df.infos<-df.infos %>% spread(feature,value)
  colnames(df.infos) <- make.names(colnames(df.infos), unique=TRUE)
  
  df.infos$LONGITUD<- as.numeric(df.infos$LONGITUD)
  df.infos$LATITUD<- as.numeric(df.infos$LATITUD)
  
  close(con)
  
  },error=function(cond) {
            message(paste("file could not be wrapped :", station_id))
  })
  return(df.infos)
}

```




```{r}
get.all.stations.infos <- function(){
  for (infos_file in all_infos_files){
  station <- strsplit(infos_file, "V")[[1]]
  station_id <- station[1]
  
  df.infos.stations <-get_stations_infos(station_id)
  
    if (exists('df.all') ){
    df.all <- bind_rows(df.all,df.infos.stations)

  } else{
    df.all <-df.infos.stations
  }
  }
  write.csv(df.all,'data/conagua_stations_infos.csv')
}

#get.all.stations.infos()


```


## Read the temperature text file for a given Station and convert the MIN/MAX Tem Prom to Dataframe

```{r}
convert_prom_temp_to_dataframe <- function(f){
  
  df.tmp.min<- as.data.frame(c(1))
  df.tmp.max<- as.data.frame(c(1))
  
  tryCatch({
 
    DATA_FOLDER <- 'data/sonaguas/'
    
    file_name <-paste(DATA_FOLDER,f,sep='')
    con <- file(file_name)
    lns <- readLines(con,encoding = 'UTF-8')
    
    MINIMA <- grep('MINIMA',lns)
    start.read.mintmp <- grep('TEMP MINIMA PROM',lns)
    start.read.maxtmp <- grep('TEMP MAXIMA PROM',lns)
    
    
    col.names <- c('title', 'YEAR',month.abb[seq(1:12)],'ACCUMULATED','MEAN', 'MONTHS')
    
    if(length(start.read.mintmp)>0){
      
      
      end.read.mintmp <- MINIMA[match(start.read.mintmp,MINIMA)+1]
       tmp.min.rows <- end.read.mintmp - start.read.mintmp -2
       
      df.tmp.min<- read.fwf(file_name, width=c(25,5,rep(7,each=13),10,10),skip=start.read.mintmp,col.names=col.names,n=tmp.min.rows,encoding = 'UTF-8')
    df.tmp.min$title='TEMP MINIMA PROM'
    }
    
     if(length(start.read.maxtmp)>0){
        end.read.maxtmp <- MINIMA[MINIMA>start.read.maxtmp][1]
          tmp.max.rows <- end.read.maxtmp - start.read.maxtmp -2
    df.tmp.max<- read.fwf(file_name, width=c(25,5,rep(7,each=13),10,10),skip=start.read.maxtmp,col.names=col.names,n=tmp.max.rows,encoding = 'UTF-8')
    df.tmp.max$title='TEMP MAXIMA PROM'
     }

    df<-  bind_rows(df.tmp.min,df.tmp.max)
   
    df$ID <- file_name
    
    df <- df %>%  dplyr::select(ID, everything())
    
    close(con)
    
  },error=function(cond) {
            message(paste("file could not be wrapped :", file_name))
  })
    
    return (df)
}



```

### Retrieve temp for stations without V_7 files (use V_3 instead)
```{r}

retrieve.all.temps.to.df <- function(){
  df<-as.data.frame(c(1))
  print (df)
  for (f in all_temp_files){
  
    df.prom.temp<- convert_prom_temp_to_dataframe(f)
    tryCatch(
      {
       df<- bind_rows(df,df.prom.temp)
      
  },error=function(cond) {
            message(paste("file could not be wrapped :", f))
  })
   
  }
  
  df<- df %>% mutate_at(vars(c(ID)) ,funs(str_extract(.,"[^/]+(?=.txt)")))
 df[1]<-NULL
write.csv(df,'data/conagua_stations_temp.csv')
}

retrieve.all.temps.to.df()
```

```{r}
df.conagua <- read.csv('data/conagua_stations_temp.csv')

```

